---
title: "Supplemental Code for AOAS1943"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Directory structure

The repo has the following directory structure.

```{r echo = FALSE}
library(listdown)
file_type <- function() {
  a = 0
  class(a) <- "file"
  a
}
dir_type <- function() {
  a = 0
  class(a) <- "dir"
  a
}

`project directory` <- list(
  "flode.Rproj" = file_type(),
  "README.Rmd" = file_type(),
  "README.md" = file_type(),
  "analysis/" = dir_type(),
  "data/" = dir_type(),
  "output/" = dir_type(),
  "preprocessing/" = dir_type()
)
ld_cc_dendro(`project directory`)
```

Each of the files and subdirectories in this repository are described below.


* `README.Rmd/md`: Readme explaining the materials in this repository
* `analysis/`
  * `202311_dataAnalysis_steps.Rmd/.html`: Code/output for processing motivating data and generating Figures 1 and 5 in the manuscript.
  * `202311_dataAnalysis_prediction.Rmd/.html`: Code/output for obtaining confidence intervals for coefficient functions shown in Figure 5 and for performing 10-fold cross validation of flode, fconc, and fhist.
  * `202312_simulation_results.Rmd/.html`: Code/output for processing simulation results and generating Figures 3 and 4 in the manuscript.
  * `202310_simulation_results_supplement.Rmd/.html`: Code/output for processing simulation results in the supplement.
* `data/20231012_tidied_steps.RDA`: Tidied mouse gait data
* `output/`: Figures, data analysis, and simulation results
  * `*.png`: Figures from the manuscript and supplement
  * `pUp_results.RData`: Data analysis results from `flode` model.
  * `pUp_fhist.RData`: Data analysis results from `fhist` model. 
  * `pUp_fconc.RData`: Data analysis results from `fonc` model. 
* `preprocessing/`: Source files for executing core modeling code, generating simulated data, and running simulations
  * `estimate_flode.R`: Defines function for estimating `flode` model using EM algorithm
  * `make_pffr.R`: Defines function for running functional historical regression model and processing surfaces for visualization
  * `search_alpha.R`: Defines function to perform grid search over possible $\alpha_0$ values to choose best initial value. 
  * `simulated_paw.R`: Defines function for simulating data.
  * `utils.R`: Defines helper functions called in the `estimate_flode.R` file.
  * `simulations/`: File for generating simulation results across all simulation scenarios.
    * `20240605_simulations_fromKalman.R`: Performs simulations from a Kalman filter model for Supplemental Figure 6.
    * `20240123_simulations_varyN.R`: Performs simulations for Supplemental Figure 4.
    * `20231212_simulations_fullComparison.R`: Generates model results for flode, fhist, and fconc when data is generated from a flode model
    * `20231212_simulations_fullComparison_genfhist.R`: Generates model results for flode, fhist, and fconc when data is generated from an fhist model
    * `20231115_simulations_confInt_boot.R`: Produces simulations for flode paper for getting coverage for bootstrap confidence intervals
    * ` 20231025_simulations_tuning.R`: Performs simulations for varying over values of flode tuning parameters for Supplemental Figure 3.


