---
title: "flode simulation results for supplement"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and creates Figures 1,3,4, and 6 from the web supplement. Libraries and code are sourced below.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(viridis)
library(splines)
library(refund)
library(Matrix)
library(mvtnorm)
library(patchwork)
library(pracma)
library(latex2exp)
#library(ggthemes)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

source(here::here("preprocessing", "simulated_paw_kalman.R"))
source(here::here("preprocessing", "predict_flode.R"))
source(here::here("preprocessing", "simulated_paw.R"))
source(here::here("preprocessing", "make_pffr.R"))
source(here::here("preprocessing", "estimate_flode.R"))
source(here::here("preprocessing", "utils.R"))
source(here::here("preprocessing", "search_alpha.R"))
```


### Computation times


```{r}
filename = here::here("output",
                      "202312_simResultsComptime.RData")

load(filename)



filename = here::here("output",
                      "202312_simResults.RData")
load(filename)
```


## Supplemental Figure 1


```{r appendix_compTimes, fig.height=4}
p1 = df %>%
  select(iter, contains("time"), alpha_true) %>%
  pivot_longer(time_flode:time_fhist, names_to = "model", values_to = "time") %>%
  mutate(time = time / 60,
         model = str_remove(model, "time_")) %>%
  filter(model != "fconc") %>%
  ggplot(aes( x = factor(alpha_true),  y = log(time), 
	            group = interaction(factor(alpha_true), model), 
	            fill = model)) +
	geom_boxplot() +
  scale_fill_brewer(type = "div",palette = 5) +
  #ylim(0, 30) +
  theme(legend.position = c(.6, .8)) +
  labs(x = TeX("True $\\alpha$"), y = "Computation time (log minutes)")


p2 = df_miter %>%
  group_by(alpha_true) %>%
  summarize(maxiter = median(miter)) %>%
  ungroup() %>%
  mutate(alpha_true = factor(alpha_true)) %>%
  ggplot(aes(group = NULL)) +
  geom_point(aes(x = alpha_true, y = maxiter)) + 
  geom_line(aes(x = as.numeric(alpha_true), y = maxiter)) +
  labs(x = TeX("True $\\alpha$"), y = "Median number of EM iterations")


p1 + p2
```




## Supplemental Figure 3

Tuning of smoothing parameters.

```{r}

results_tuning = c()
files = list.files(here::here("output", "simulation_results"), 
									 pattern = "flode_tuning",
									 full.names =  TRUE)

for(f in 1:length(files)){
	load(files[f])
	results_tuning = rbind(results_tuning, as_tibble(results))
}
```



and plots for beta est

```{r}
p_tune = results_tuning %>%
  filter(alpha_true == 4) %>%
  mutate(Kt = factor(Kt, labels = c("K = 10", "K = 20", "K = 40")),
         alpha_true = str_c("alpha = ", alpha_true),
         a = str_c("lambda = ", a)) %>%
  ggplot(aes(Kt, y = surfaceErr_flode, 
             fill = Kt)) +
  geom_boxplot() +
  scale_fill_brewer(type = "div",palette = 1) +
  theme(legend.position = "none", strip.text = element_text(size = 14), axis.title.x = element_blank()) +
  facet_wrap(~a) +
  labs(y = "flode surface error") 

```


One data set where alpha =4, Kt = 40, a = 0.001

```{r, cache = TRUE}
simulated_data = simulate_flode(I = 100, 
																		D = 50, 
																		sigma_y0 = 5, 
																		sigma = 0.1, 
																		alpha = 4, 
																		rand_int = TRUE, 
																		lambda0 = 10,
																		seed = 335)

dat = simulated_data$data %>% mutate(int = 1)

flode_results_a001 = estimate_flode(dat, 
																	 Kt = 40, 
																	 alpha0 = 4,  
																	 max_iter = 200, 
																	 forcing_functions = c("int", "x"),
																	 tol =  0.00001, 
																	 sigma_b = 20,
																	 sigma_d = 20,
																	 random_effects = TRUE,
																	 initial_position = TRUE,
																	 y0_vec = simulated_data$y0,
																	 a = 0.001)
	
flode_results_a01 = estimate_flode(dat, 
																	 Kt = 40, 
																	 alpha0 = 4,  
																	 max_iter = 200, 
																	 forcing_functions = c("int", "x"),
																	 tol =  0.00001, 
																	 sigma_b = 20,
																	 sigma_d = 20,
																	 random_effects = TRUE,
																	 initial_position = TRUE,
																	 y0_vec = simulated_data$y0,
																	 a = 0.01)
```


Further simulate scenarios for alpha =4, Kt = 40, both a values, and do


```{r}
ptune2 = simulated_data$coefficient_fns %>%
  select(time, truth = beta1) %>%
  mutate(a001 = flode_results_a001$beta$beta1,
         a01 = flode_results_a01$beta$beta1) %>%
  pivot_longer(truth:a01, names_to = "model", values_to = "beta1") %>%
  mutate(model = factor(model, levels = c("truth", "a01", "a001"), 
                        labels = c("truth", "lambda = 0.01", "lambda = 0.001"))) %>%
  ggplot(aes(time, beta1, group = model, color = model, linetype = model)) +
  geom_line(size = 1.25) +
  labs(y = TeX("$\\beta_1(t)$")) +
  theme(legend.position = c(.35, .2), legend.title = element_blank(), legend.text = element_text(size = 12))
  
```




```{r appendix_tuning,   fig.width = 9, fig.height = 4}
p_tune + ptune2
```

## Supplemental Figure 5

Figure of alpha and beta results across different trials sizes (with)

```{r}

filename = here::here("output",
                      "202401_simVaryN.RData")

load(filename)
```


Plot the results in the same manner as Figure 3 in the paper

```{r appendix_varyN, fig.height = 7}
df_varyN %>%
  mutate(alpha_bias = alpha_true - alpha_hat) %>%
  select(iter, N, alpha_true, alpha_bias, ISEb0, ISEb1) %>%
  pivot_longer(alpha_bias:ISEb1, names_to = "measure", 
               values_to = "value") %>%
  mutate(measure = factor(measure, 
                          levels = c("alpha_bias", "ISEb0", "ISEb1"),
                          labels = c("alpha-hat(alpha)", TeX("$\\int_t\\{\\beta_0(t)-\\hat{\\beta}_0(t)\\}dt$"), TeX("$\\int_t\\{\\beta_1(t)-\\hat{\\beta}_1(t)\\}dt$"))),
         N = factor(N)) %>%
  mutate(alpha = factor(alpha_true,
                        levels = c(0.5, 4, 12),
                        labels = c("alpha: 0.5", "alpha: 4",
                                   "alpha: 12"))) %>%
  ggplot(aes(N, value, group = N)) +
  geom_violin() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = "N", y = "error") +
  facet_grid(alpha~measure, scales = "free",
             labeller = label_parsed)

```



## Supplemental Figure 6

Shows results of doing prediction for simulations where data is generated from a kalman filter


First do prediction
 
```{r}
filename_k = here::here("output",
                      "202406_simResults_predictionKalman.RData")

if(file.exists(filename_k)){
	load(filename_k)
}else{
  filename = paste0("/Users/JWROBE8/onedrive/Data/flode/kalman")
  files = list.files(filename, 
                   full.names =  TRUE)
   ## generate list of datasets
  pred_df = simulate_kalman(I = 10000, 
                                 D = 50, 
                                 sigma = 0.1, 
                                 rand_int = TRUE, 
                                 lambda0 = 10,
                                 seed = 2352352)

	
  df_pred_kalman = map_dfr(files, get_predictions_genKalman, new_data = pred_df)
  
  save(df_pred_kalman, file = filename_k)
}
```



Plot the results

```{r appendix_kalman,fig.width = 2, fig.height = 3}
df_pred_kalman %>%
  select(iter, contains("mae")) %>%
  pivot_longer(flode_mae:fhist_mae, names_to = "model", 
               values_to = "MAPE") %>%
  mutate(model = str_remove(model, "_mae"),
         logMAPE = log(log(MAPE)),
         model = factor(model, levels = c("flode", "fhist"))) %>%
  ggplot(aes(model, logMAPE, group = model, fill = model)) +
  geom_boxplot() +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  labs(y = "log MAPE") +
  scale_fill_brewer(drop = FALSE, type = "div") 

```

