---
title: "Data analysis figures and results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads produces figures and runs analysis for the motivating data. First we load libraries and source files for estimating the flode model.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(viridis)
library(splines)
library(refund)
library(Matrix)
library(mvtnorm)
library(patchwork)
library(pracma)
library(tictoc)
library(vbvs.concurrent) # devtools::install_github("jeff-goldsmith/vbvs.concurrent")
library(fields)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

source(here::here("preprocessing", "estimate_flode.R"))
source(here::here("preprocessing", "utils.R"))
source(here::here("preprocessing", "search_alpha.R"))
```

# Load data

Next we load the cleaned data.

```{r data_import}
load(here::here("data", "20231012_tidied_steps.RDA"))
```



# Analysis



## data organization

For our analysis we focused on running the model on the up (`p_up`) direction, so we clean the data and focus on that variable.


```{r}
reseg = reseg %>%
  mutate(int = 1) %>%
  nest(data = c(-step)) %>%
  mutate(trial = row_number()) %>% 
  unnest(data) %>%
  select(trial, int, time = index, 
         value = p_up, v_up, bicep, tricep) %>%
  group_by(trial) %>%
  mutate(y0 = first(value))  %>%
  ungroup() 
  

```



## Modeling

First we do a grid search to find the best initial value of $\alpha$. For the grid search we run the model without random effects, choosing a particular value of $\alpha$ and using the observed initial positions to calculate the value of the loss function at that particular $\alpha$.

To reduce computation time, we run flode on the data and save flode data object. The code below only runs the flode analysis if the saved `pUp_results.RData` object doesn't already exist in the `output` folder.


### flode


```{r loss}
	alpha_start_pUp = search_alpha(data = reseg, 
														 forcing_functions = c("int",  "bicep", "tricep"),
														 Kt = 20, 
														 alpha_min = 0,
														 alpha_max = 20)
```


```{r pUp}
filename = here::here("output", "pUp_results.RData")

if(file.exists(filename)){
	load(filename)
}else{
	tic()
	# initiate alpha
	alpha_start_pUp = search_alpha(data = reseg, 
														 forcing_functions = c("int",  "bicep", "tricep"),
														 Kt = 20, 
														 alpha_min = 0,
														 alpha_max = 6)
	
	# run flode
	pUp_results = estimate_flode(reseg, 
															alpha0 = alpha_start_pUp$best_alpha, 
                            	Kt = 20,
															forcing_functions = c("int", "bicep","tricep"),
															max_iter = 50, 
															tol = 0.001, 
															initial_position = TRUE,
															random_effects = TRUE,
															alpha_upper = 6)
	time_pUp = toc()

	save(pUp_results, time_pUp, alpha_start_pUp, file = filename)
}


```


### fhist

Next we perform analysis using the functional historical model (fhist).


```{r fhist}
filename = here::here("output", "pUp_fhist.RData")


if(file.exists(filename)){
	load(filename)
}else{
  source(here::here("preprocessing", "make_pffr_pup.R"))
	tic(quiet = TRUE)
	pUp_fhist = make_pffr_pUp(reseg, covar = c("bicep","tricep"),
	                      random_int = TRUE)
	time_fhist = toc()


	save(pUp_fhist, time_fhist, file = filename)
}

```


### functional concurrent model

Finally, we perform analysis using the functional linear concurrent model (fconc).

```{r fconc}
filename = here::here("output", "pUp_fconc.RData")

if(file.exists(filename)){
	load(filename)
}else{

	tic(quiet = TRUE)
	pUp_fconc = vb_concurrent(value ~ int + bicep + tricep | time, 
	                          id.var = "trial",
                            data = reseg, 
	                          Kt = 20,
	                          Kp = 15,
                            t.min = 0, t.max = 1, 
                            standardized = TRUE)
	time_fconc = toc()


	save(pUp_fconc, time_fconc, file = filename)
}


```



# Results



## Data summary

* Number of trials is `r length(unique(reseg$trial))`
* Length of each trial is `r length(unique(reseg$time))`
* alpha is `r pUp_results$alpha`


## Computation times

* flode is `r (time_pUp$toc - time_pUp$tic)/60` minutes 
* functional historical model is `r (time_fhist$toc - time_fhist$tic)/60` minutes
* functional concurrent model is `r (time_fconc$toc - time_fconc$tic)` seconds


## Data reorg

Here we do some processing on the objects returned from the three models (flode, fhist, fconc) to calculate quantities used for comparing results across models and for modeling.

```{r}
fhist_df = pUp_fhist$fhist_df
pUp_fhist = pUp_fhist$mod


source(here::here("preprocessing", "simulated_paw.R"))
intercept_surface = make_sim_surface(alpha = pUp_results$alpha, 
                                     beta = pUp_results$beta$beta0, 
                                     D = 100, t_max = 1)

# beta coefficients (doesn't include alpha term)
beta_df = pUp_results$beta %>%
  rename(bicep_flode = beta1,
         tricep_flode = beta2) %>%
  mutate(beta0_flode = as.numeric(rowSums(intercept_surface))/100,
         beta0_fconc = pUp_fconc$beta.pm$int,
         bicep_fconc = pUp_fconc$beta.pm$bicep,
         tricep_fconc = pUp_fconc$beta.pm$tricep,
         beta0_fhist = as.numeric(coef(pUp_fhist, n1 = 100, n2 = 100)$smterms$"Intercept(t)"$coef$value),
         bicep_fhist = NA,
         tricep_fhist = NA) %>%
  select(time, everything())
 


# beta surfaces (includes alpha term)
make_surface_flode = function(data, beta_num){
  surface = data$surface[[beta_num]]
  surface %>% as_tibble() %>%
    mutate(t = attributes(surface)$t) %>%
    gather(s, value, starts_with("s")) %>%
    arrange(t) %>%
    mutate(s = rep(attributes(surface)$s, dim(data$surface[[1]])[1]),
           label = paste0("beta", beta_num - 1))
}

surface_flode = map_dfr(2:3, make_surface_flode, data = pUp_results)

# fhist surfaces
make_surface_fhist = function(fhist_surf, D = 100, label){
		## interpolate surface to get on same grid as flode
		interp_obj = list(x = unique(fhist_surf$s), y = unique(fhist_surf$t),
											z = matrix(fhist_surf$value, D, D, byrow = TRUE))
		
		t = s = seq(0, 1, length.out = D)
		loc_mat = list(x = s, y = t)
		surface = fields::interp.surface.grid(interp_obj, loc_mat)$z
		surface[D,] = surface[D-1,] # carry last value forward so not NA
		surface =	surface * lower.tri(surface, diag = TRUE) %>% as_tibble() 


		surface %>% 
		  mutate(t = t) %>%
		  pivot_longer(V1:V100, names_to = "s", values_to = "value") %>%
		  mutate(label = label,
		         s = rep(seq(0, 1, length.out = D), 100)) 

}

fhist_surf1 = coef(pUp_fhist, n1 = 100, n2 = 100)$smterms$"ff(x1,s)"$coef %>%
  select(s = x1.smat, t = x1.tmat, value = value) %>% as_tibble() 

fhist_surf2 = coef(pUp_fhist, n1 = 100, n2 = 100)$smterms$"ff(x2,s)"$coef %>%
  select(s = x2.smat, t = x2.tmat, value = value) %>% as_tibble() 

		
surface_fhist = bind_rows(make_surface_fhist(fhist_surf1, label = "bicep_fhist"),
                          make_surface_fhist(fhist_surf2, label = "tricep_fhist")) %>%
  select(value_fhist = value)


surface_flode = surface_flode %>%
  mutate(label = ifelse(label == "beta1", "bicep","tricep"))


surfaces = bind_cols(surface_flode, surface_fhist) %>%
  pivot_longer(c(value, value_fhist), names_to = "model", values_to = "value") %>%
  mutate(model = ifelse(model == "value", "flode", "fhist"))
  
```



Now, get xB for the fhist model.


```{r}
integrate_surface = function(basis_time, surface, covar){
  D = length(basis_time)
  
  beta_covar_f = function(s0, t0 = 1, basis_time, surface, covar){
    filter(surface, s == s0, t == t0)$value * covar[which(basis_time == s0)] 
  }
  
  Gstar = sapply(basis_time, FUN = function(t){cumtrapz(basis_time, beta_covar_f(basis_time, t, basis_time, surface,covar))[D,]})
  return(t(Gstar))
}

pUp_fhist
xB_fhist = matrix(NA, length(unique(reseg$trial)), 100)
grid = sort(unique(reseg$time))
surface_bicep = make_surface_fhist(fhist_surf1, label = "bicep_fhist")
surface_tricep = make_surface_fhist(fhist_surf2, label = "tricep_fhist")
b0 = coef(pUp_fhist)$pterms[1]
for(i in unique(reseg$trial)){
  bicep = filter(reseg, trial == i)$bicep
  tricep = filter(reseg, trial == i)$tricep
  xb_bicep = integrate_surface(grid, surface_bicep, covar = bicep) 
  xb_tricep = integrate_surface(grid, surface_tricep, covar = tricep) 
  
  xB_fhist[i,] = b0 +beta_df$beta0_fhist +
    xb_bicep + xb_tricep
}

rm(grid, surface_bicep, surface_tricep, i, xb_bicep, xb_tricep, bicep, tricep)
```






```{r}
# fitted values, linear predictor without random effects, random effects 
df_results = pUp_results$data %>% 
  mutate(xB = pUp_results$xB,
         lp_flode = xB + y0_star,
         lp_fconc = pUp_fconc$fixef,# fixef for 
         lp_fhist = as.vector(t(xB_fhist)),
         yhat_fhist = as.vector(t(fitted(pUp_fhist))),
         yhat_fconc = pUp_fconc$Yhat,
         resid_flode = value - lp_flode,
         resid_fconc = value - lp_fconc,
         resid_fhist = value - lp_fhist) %>%
  rename(yhat_flode = yhat)


cv_df <- readRDS(df, file = here::here("output", "predict_fold_all.rds")) %>%
  pivot_longer(yhat_flode:yhat_fconc, names_to = "method", values_to = "yhat",
               names_prefix = "yhat_") %>%
  mutate(resid = value-yhat,
         mae = abs(resid)) %>%
  group_by(time, method) %>%
  summarize(mae = mean(mae)) %>%
  ungroup() %>%
  arrange(method, time)
```






```{r}
rm(list=lsf.str())
```


## Plots

### Figure 1

This produces Figure 1 in the manuscript, and saves it to the `output` folder as `fig_steps-1.png`.


```{r fig_steps, fig.height = 5.5, fig.width = 9}
set.seed(32543)
steps = sample(unique(reseg$trial), 5)
reseg %>%
  filter(trial %in% steps) %>%
  pivot_longer(value:tricep, 
               names_to = "var",
               values_to = "value") %>%
  mutate(var =  factor(var, levels = c("tricep", "bicep", 
                                       "v_up", "value"),
                       labels = c("triceps activity", 
                                  "biceps activity", 
                                  "velocity",
                                  "position"))) %>%
  ggplot(aes(time, value)) +
  geom_line() +
  facet_grid(var~trial, scales = "free_y", 
             labeller = labeller(var = label_value,
                                 trial = label_both)) +
  theme(strip.text = element_text(size = 12), 
        #strip.text = element_blank(1),
        axis.title.x = element_blank()) #

df = reseg %>%
  mutate(trial = factor(trial)) %>%
  pivot_longer(value:tricep, names_to = "var", values_to = "value") %>%
  mutate(var =  factor(var, levels = c("tricep", "bicep", "v_up", "value"),
                       labels = c("triceps activity", 
                                  "biceps activity", "velocity",
                                  "position")))

```


### Figure 5

This produces Figure 5 in the manuscript, and saves it to the `output` folder as `fig_wald_boot-1.png`. The Figure is created by modifying a figure with flode coefficient functions and 95% confidence intervals called `flode_coefs_plot` that is created in the `202311_dataAnalysis_prediction.Rmd` file.


```{r fig_wald_boot, fig.height = 5, fig.width = 10}
load(file = here::here("output", "flode_coefs_plot.Rda"))


p2 = df %>% 
  ggplot(aes(time, value, group = trial)) + 
  geom_line(alpha = 0.2) +
  facet_wrap(~var, nrow = 1, scales = "free_y") + 
  theme(strip.text = element_text(size = 14),
        axis.title.x = element_blank())


p2/flode_coefs_plot


```




# Supplement

Here we provide code to reproduce the supplemental figures that use the real data.

### Figure S5

Surfaces plots for flode and fhist

```{r}

make_surface_df = function(data, beta_num, model = "flode", D = 100){
  if(model == "flode"){
      surface = data$surface[[beta_num]]
  
  
  surface %>% as_tibble() %>%
    mutate(t = seq(0,1, length.out = D)) %>%
    gather(s, value, starts_with("s")) %>%
    arrange(t) %>%
    mutate(s = rep(seq(0,1, length.out = D), dim(data$surface[[1]])[1]),
           label = paste0("beta", beta_num - 1),
           model = model)
  }
  else if(model == "fhist"){
     ## calculate surface for fhist
    if(beta_num == 1){
      fhist_surface = coef(data, n1 = D, n2 = D)$smterms$"ff(x1,s)"$coef %>%
         select(s = x1.smat, t = x1.tmat, value = value) %>% 
        as_tibble()
        
    }else if(beta_num == 2){
      fhist_surface = coef(data, n1 = D, n2 = D)$smterms$"ff(x2,s)"$coef %>%
         select(s = x2.smat, t = x2.tmat, value = value) %>% 
        as_tibble()
    }
    
    fhist_surface = fhist_surface %>%
      select(t, s, value)
  
  ## interpolate surface to get on same grid as flode
  interp_obj = list(x = unique(fhist_surface$s), y = unique(fhist_surface$t),
                    z = matrix(fhist_surface$value, D, D, byrow = TRUE))
  
  tt = ss = seq(0, 1, length.out = D)
  loc_mat = list(x = ss, y = tt)
  fhist_surface = fields::interp.surface.grid(interp_obj, loc_mat)$z
  fhist_surface[D,] = fhist_surface[D-1,] # carry last value forward so not NA
  fhist_surface = fhist_surface * lower.tri(fhist_surface, diag = TRUE)# zero out upper triangle of the matrix
  
  colnames(fhist_surface) = paste0("s", ss)
  fhist_surface = fhist_surface %>%
    as_tibble() %>%
    mutate(t = tt) %>% select(t, everything()) %>%
    gather(s, value, starts_with("s")) %>%
    arrange(t) %>%
    mutate(s = rep(ss, D),
           label = paste0("beta", beta_num),
           model = model,
           value = ifelse(s <= t, value, NA))
  
  fhist_surface
  }
}

surface_df = map_dfr(2:3, make_surface_df, data = pUp_results)
surface_df_fhist = map_dfr(1:2, make_surface_df, data = pUp_fhist, model = "fhist")

surface_df = bind_rows(surface_df, surface_df_fhist)
```

Saved as `appendix_surfaces-1` in the `output` folder.


```{r appendix_surfaces}

surface_df %>%
  mutate(coef = factor(label, levels = c("beta2", "beta1"), 
                       labels = c("triceps", "biceps")),
         value = ifelse(s>t, NA, value)) %>%
  ggplot(aes(s, t)) + 
  geom_tile(aes(fill = value, col = value)) + 
  scale_fill_viridis_c() +
  scale_colour_viridis_c() +
  facet_grid(model ~ coef) +
  theme(legend.position = "bottom")

```
