---
title: "Data analysis figures and results for steps data"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads produces Figures and runs analysis using the real data.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(stringr)
library(viridis)
library(splines)
library(refund)
library(Matrix)
library(mvtnorm)
library(patchwork)
library(pracma)
library(tictoc)
library(vbvs.concurrent)
library(fields)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

source(here::here("preprocessing", "estimate_flode.R"))
source(here::here("preprocessing", "predict_flode.R"))
source(here::here("preprocessing", "utils.R"))
source(here::here("preprocessing", "search_alpha.R"))
source(here::here("preprocessing", "make_pffr_pup.R"))
```

# Load paw data

```{r data_import}
load(here::here("data", "20231012_tidied_steps.RDA"))
```



# Analysis



## data organization


```{r}
reseg = reseg %>%
  mutate(int = 1) %>%
  #filter(step <= 208) %>%
  nest(data = c(-step)) %>%
  mutate(trial = row_number()) %>% 
  unnest(data) %>%
  select(trial, int, time = index, 
         value = p_up, v_up, bicep, tricep) %>%
  group_by(trial) %>%
  mutate(y0 = first(value))  %>%
  ungroup() 
  

```



```{r}
# simulated_data = simulate_flode(I = 25, 
# 																		D = 40, 
# 																		sigma_y0 = 5, 
# 																		sigma = 0.1, 
# 																		alpha = 3, 
# 																		rand_int = TRUE, 
# 																		lambda0 = 20,
# 																		seed = 12345)
# 
# reseg = simulated_data$data %>% 
#   mutate(int = 1, bicep = x, 
#          tricep = sin(time) + rnorm(1))
```


## Splitting data

Here we run 10-fold cross validation with flode, fhist, and fconc to determine which model has the best predictive performance.

```{r}
# set up the folds
set.seed(22334)
folds = sample(1:10, size = length(unique(reseg$trial)),
       replace = TRUE)

reseg = reseg %>%
  nest(data = c(-trial)) %>%
  mutate(fold = folds) %>%
  unnest(data)
```

## Cross-validation

Set up functions for integrating fhist model.

```{r}
# fhist surfaces
make_surface_fhist = function(fhist_surf, D = 100, label = ""){
		## interpolate surface to get on same grid as flode
		interp_obj = list(x = unique(fhist_surf$s), y = unique(fhist_surf$t),
											z = matrix(fhist_surf$value, D, D, byrow = TRUE))
		
		t = s = seq(0, 1, length.out = D)
		loc_mat = list(x = s, y = t)
		surface = fields::interp.surface.grid(interp_obj, loc_mat)$z
		surface[D,] = surface[D-1,] # carry last value forward so not NA
		surface =	surface * lower.tri(surface, diag = TRUE) %>% as_tibble() 


		surface %>% 
		  mutate(t = t) %>%
		  pivot_longer(V1:V100, names_to = "s", values_to = "value") %>%
		  mutate(label = label,
		         s = rep(seq(0, 1, length.out = D), 100)) 

}

integrate_surface = function(basis_time, surface, covar){
  D = length(basis_time)
  
  beta_covar_f = function(s0, t0 = 1, basis_time, surface, covar){
    filter(surface, s == s0, t == t0)$value * covar[which(basis_time == s0)] 
  }
  
  Gstar = sapply(basis_time, FUN = function(t){cumtrapz(basis_time, beta_covar_f(basis_time, t, basis_time, surface,covar))[D,]})
  return(t(Gstar))
}

```


Set up function for doing cross-validation.

```{r}
do_cv = function(fold_num){
    train = filter(reseg, fold != fold_num)
    test = filter(reseg, fold == fold_num)
  
    predict_df = select(test, time, value, fold)
  ######################################################################
  ## run flode
  ######################################################################
    alpha_start_pUp = search_alpha(data = train, 
														 forcing_functions = c("int",  "bicep", "tricep"),
														 Kt = 20, 
														 alpha_min = 1,
														 alpha_max = 8)
	
	# run flode
	flode_results = estimate_flode(train, 
															alpha0 = alpha_start_pUp$best_alpha, 
                            	Kt = 20,
															forcing_functions = c("int", "bicep","tricep"),
															max_iter = 50, 
															tol = 0.001, 
															initial_position = FALSE,
															random_effects = TRUE)
	
	
	# calculate cross-validated error
	# also store alpha for each around as well as testing fold
	predict_df$yhat_flode = predict_flode(flode_results, 
	                                      test,
	                                      forcing_functions = c("int", "bicep","tricep"))$yhat_flode
	
	######################################################################
  ## run fhist
  ######################################################################
	
	fhist_results = make_pffr_pUp(train, covar = c("bicep","tricep"),
	                          random_int = TRUE)
	
	
	xB_fhist = xB_fconc = matrix(NA, length(unique(test$trial)), 100)
	grid = sort(unique(train$time))
	
	fhist_surf1 = coef(fhist_results$mod, n1 = 100, n2 = 100)$smterms$"ff(x1,s)"$coef %>%
	  select(s = x1.smat, t = x1.tmat, value = value) %>% as_tibble() 
	
	fhist_surf2 = coef(fhist_results$mod, n1 = 100, n2 = 100)$smterms$"ff(x2,s)"$coef %>%
	  select(s = x2.smat, t = x2.tmat, value = value) %>% as_tibble() 
	
	surface_bicep = make_surface_fhist(fhist_surf1, label = "bicep_fhist")
	surface_tricep = make_surface_fhist(fhist_surf2, label = "tricep_fhist")
	b0 = coef(fhist_results$mod)$pterms[1]
	beta0_fhist = as.numeric(coef(fhist_results$mod, n1 = 100, n2 = 100)$smterms$"Intercept(t)"$coef$value)
	
	trials = unique(test$trial)
	for(i in 1:length(trials)){
	  bicep = filter(test, trial == trials[i])$bicep
    tricep = filter(test, trial == trials[i])$tricep
    xb_bicep = integrate_surface(grid, surface_bicep, covar = bicep) 
    xb_tricep = integrate_surface(grid, surface_tricep, covar = tricep) 
  
    xB_fhist[i,] = b0 + beta0_fhist +
      xb_bicep + xb_tricep
    }
	
	predict_df$yhat_fhist = as.vector(t(xB_fhist))
	
	######################################################################
  ## run fconc
  ######################################################################
	# 5 has an error
	fconc_results = vb_concurrent(value ~ int + bicep + tricep | time, 
	                              id.var = "trial",
	                              data = train, 
	                              Kt = 20,
	                              Kp = 15,
	                              t.min = 0, t.max = 1, 
	                              standardized = TRUE)
	
	
		trials = unique(test$trial)
		for(i in 1:length(trials)){
		  bicep = filter(test, trial == trials[i])$bicep * fconc_results$beta.pm$bicep
      tricep = filter(test, trial == trials[i])$tricep * fconc_results$beta.pm$tricep
      
      xB_fconc[i,] = fconc_results$beta.pm$int + bicep + tricep
		}
		
		predict_df$yhat_fconc = as.vector(t(xB_fconc))
	
	predict_df
}# end function


```

Perform cross validation

```{r}
filename = here::here("output", "predict_fold_1.rds")

if(file.exists(filename)){
	print("not rerunning prediction")
}else{
  #predict_df = map_dfr(1:5, do_cv)
  
  for(fold in 1:10){
    predict_df = do_cv(fold)
    
    f = paste0("predict_fold_", fold, ".rds")
    saveRDS(predict_df, file = here::here("output", f))
  }
 }
```



load data from each fold and process into table and figure.

```{r}

f1 = readRDS(here::here("output", "predict_fold_1.rds"))
f2 = readRDS(here::here("output", "predict_fold_2.rds"))
f3 = readRDS(here::here("output", "predict_fold_3.rds"))
f4 = readRDS(here::here("output", "predict_fold_4.rds"))
f5 = readRDS(here::here("output", "predict_fold_5.rds"))
f6 = readRDS(here::here("output", "predict_fold_6.rds"))
f7 = readRDS(here::here("output", "predict_fold_7.rds"))
f8 = readRDS(here::here("output", "predict_fold_8.rds"))
f9 = readRDS(here::here("output", "predict_fold_9.rds"))
f10 = readRDS(here::here("output", "predict_fold_10.rds"))


df = bind_rows(f1, f2, f3, f4, f5, 
               f6, f7, f8, f9, f10)


saveRDS(df, file = here::here("output", "predict_fold_all.rds"))
```



The following table shows mean absoute error for each model. These numbers appear as MAE in the text of the paper.

```{r}

df %>%
  pivot_longer(yhat_flode:yhat_fconc, names_to = "method", values_to = "yhat",
               names_prefix = "yhat_") %>%
  mutate(resid = value-yhat,
         mse = resid^2,
         mae = abs(resid)) %>%
  group_by(method) %>%
  summarize(mae = mean(mae)) %>%
  ungroup() %>%
  arrange(mae) %>%
  knitr::kable(digits = 3)

```


# Bootstrap

Get bootstrap confidence intervals for $\beta(t)$. 

## Generate bootstrap samples

First, generate bootstrap sample of the full data.

```{r}
# set up the folds
set.seed(22334)
reseg = reseg %>%
  select(trial, int, time, value, bicep, tricep, y0) %>%
  nest(data = c(-trial))


reseg_boot <- bootstraps(reseg,
                          times = 1000)
```

For our inference procedure, we first fit `flode` to obtain an estimated value of $\alpha$, then generate $n_b$ bootstrap samples and refit flode to obtain bootstrap estimates for the rest of the parameters.

```{r}
# load full model
filename = here::here("output", "pUp_results.RData")
load(filename)

bootname = here::here("output", "pUp_results_boot.rda")

if(file.exists(bootname)){
	load(bootname)
}else{
  reseg_betas <- reseg_boot %>%
  mutate(beta = map(splits, ~predict_flode(flode_model = pUp_results,
                                            new_data = .,
                                            forcing_functions = c("int",  "bicep", "tricep"),
                                            boot = TRUE) )
         )


 save(reseg_betas, file = bootname)
 }

```


Using our bootstrapped estimates, we create confidence intervals for the coefficient functions. We then plot the coefficient function and 95% confidence intervals and sae this plot as `flode_coefs_plot.Rda`. This plot is further modified in the `202311_dataAnalysis_steps.Rmd` file and becomes part of Figure 5 in the manuscript.

```{r}

beta_full = pUp_results$beta %>%
  pivot_longer(beta0:beta2, names_to = "coef", values_to = "value") %>% filter(coef != "beta0")

reseg_betas_pl = reseg_betas %>%
  unnest(beta) %>%
  select(-beta0) %>%
  pivot_longer(beta1:beta2, names_to = "coef", values_to = "value") %>%
  group_by(time, coef) %>%
  summarize(sd = sd(value)) %>%
  ungroup() %>%
  mutate(beta_hat = beta_full$value,
         lower = beta_hat - 2 * sd,
         upper = beta_hat + 2 * sd) %>%
    mutate(coef = factor(coef, levels = c("beta2", "beta1"), 
                       labels = c("beta[triceps](t)", "beta[biceps](t)"))) 

flode_coefs_plot = reseg_betas_pl %>%
  ggplot() +
  geom_hline(yintercept = 0, color = "red") +
  geom_line(aes(time, lower), linetype = 2) +
  geom_line(aes(time, upper), linetype = 2) +
  geom_line(aes(time, beta_hat)) +
  labs(y = "value") +
  theme(strip.text = element_text(size = 14)) +
  facet_wrap(~coef, scales = "free",
             labeller = label_parsed)

save(flode_coefs_plot, file = here::here("output", "flode_coefs_plot.Rda"))
```

